---
title: "Mastering shiny"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
```


## Ch1: My first Shiny app

### Exercise 1

This is the most basic Shiny App. It contains `UI` and a `Server` elements. 

```{r}
ui <- fluidPage(
  "Hello, world!"
)
server <- function(input, output, session) {
}
shinyApp(ui, server)
```



### Exercise 2

For th `UI`, we add some elements in the panel

```{r}
ui <- fluidPage(
  selectInput("dataset", label="Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)
server <- function(input, output, session) {
}
shinyApp(ui, server)
```

## Exercise 3

Adding functions to thee `server` to plot the 

```{r}
ui <- fluidPage(
  selectInput("dataset", label="Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)
server <- function(input, output, session) {
  
  # Output para la el 'verbatim' summary
  output$summary <- renderPrint({
    dataset <- get(input$dataset, "package:datasets")
    summary(dataset)
  })
  
  # Output para la table
  output$table <- renderTable({
    dataset <- get(input$dataset, "package:datasets")
    dataset
  })
}
shinyApp(ui, server)
```


### Exercise 4: Optimizing with `Reactive` expresions

Giving the fact that we duplicate the `get(input$dataset, "package:datasets")` expression, we can use a variable to do this beforehand.


```{r}
ui <- fluidPage(
  selectInput("dataset", label="Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)
server <- function(input, output, session) {
  
  # Reactive expression for the input
  
  dataset <- reactive({
    get(input$dataset, "package:datasets")
  })
  
  # Output para la el 'verbatim' summary
  output$summary <- renderPrint({
    summary(dataset())
  })
  
  # Output para la table
  output$table <- renderTable({
    dataset()
  })
}
shinyApp(ui, server)
```


## Ch2: Basic UI

### Inputs

The inputs have:

1. `inputID`: Single name that connect with the `server`. It must be unique
2. `label`: Create a human-readable label for the control
3. `value`: the value of that input. 

```{r}

# Limited choices
  animals <- c("dog", "cat", "mouse", "otro")


ui <- fluidPage(
  selectInput("dataset", label="Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table"),
  
  # Type of Inputs
  ## Slider
  sliderInput("min", "Limit (minimum)", value = 50, min = 0, max = 100),
  
  ## Free text
  textInput("name", "What's your name?"),
  passwordInput("password", "What's your password?"),
  textAreaInput("story", "Tell me about yourself", row=3),
  
  
  # Numeric Inputs
  numericInput("num", "Number one", value=0, min=0, max=100),
  sliderInput("num2", "Number two", value=0, min=0, max=100),
  sliderInput("rang", "Range of values", value = c(10,20), min = 0, max = 100),
  
  # Dates
  dateInput("Date", "when you born?"),
  dateRangeInput("Holiday", "When do you want go to vacation?"),
  
  
  # Limited choices
  selectInput("statew", "What is you favorite state", choices=state.name),
  radioButtons("animal", "What's your favorite animal?", choices = animals),
  radioButtons("faces", "Choose one face?",
               choiceNames = list(
                 fontawesome::fa(name = "r-project", fill = "steelblue"),
                  icon("sad-tear")
               ),
               choiceValues = list(
                 "Fontawsome", "icon"
               )),
  
  # Dropdowns
  selectInput("dropdown","what is your favorite state", choices = state.name, multiple = TRUE),
 
  # Checkbox
  checkboxGroupInput("animals_check", "What animals do you like?", choices=animals),
  
  checkboxInput("clean", "Limpiamos?", value = TRUE),
  checkboxInput("shutdown", "No shutdown"),
  
  # File Uploads
  fileInput("upload_data", NULL),
  
  # Actions buttos
  actionButton("click", "Clilck me!!", class = "btn-danger"),
  actionButton("Click2", "Drink me", 
               icon =  icon("cocktail"))
)

  

  server <- function(input, output, session) {
  
  # Reactive expression for the input
  
  dataset <- reactive({
    get(input$dataset, "package:datasets")
  })
  
  # Output para la el 'verbatim' summary
  output$summary <- renderPrint({
    summary(dataset())
  })
  
  # Output para la table
  output$table <- renderTable({
    dataset()
  })
}
shinyApp(ui, server)
```


### Outputs

Each `output` function on the front is coupled with a `render`

There are three main outputs:
1. `Text`: 
2. `Table`: 
3. `Plots`: 

```{r}

# Limited choices
  animals <- c("dog", "cat", "mouse", "otro")


ui <- fluidPage(
  # Outputs
  ## Text
  textOutput("Text"),
  verbatimTextOutput("code"),
  
  
  ## Tables
  tableOutput("static"),
  dataTableOutput("dynamic"),
  
  #Plots
  plotOutput("plot", width = "400px"),
  
  # Download
  downloadButton()
)

  

  server <- function(input, output, session) {
  
  # Reactive expression for the input
  
  dataset <- reactive({
    get(input$dataset, "package:datasets")
  })
  
  # Output para la el 'verbatim' summary
  output$Text <- renderText({
    "Hello, I am testing renderText"
  })
  
  # Output para la table
  output$code <- renderPrint({
    summary(1:10)
  })
  
  
  ## Testing thee difference
  output$Text <- renderText("Hello friend!")
  output$code <- renderPrint(summary(1:10))
  
  
  # Tables
  ## Static
  output$static <- renderTable(head(mtcars))
  ## Dynamic
  output$dynamic <- renderDataTable(mtcars, options = list(pageLength=5))
  
  
  # Plots
  output$plot <- renderPlot(plot(1:5), res=100)
  
  # Download
}
shinyApp(ui, server)

```




## Ch2: Basic Reactivity


### Input

```{r}

ui <- fluidPage(
  # front end interface
  numericInput("count", label="Number of values", value = 100)
)

server <- function(input, output, session) {
  # back end logic
   message("The value of input$count is ", input$count)
}

shinyApp(ui, server)
```

### Output

```{r}
ui <- fluidPage(
  textOutput("greeting")
)

server <- function(input, output, session) {
  output$greeting <- renderText("Hello human!")
  
  output$greeting <- "Hello human"  # Error: You forget the render function.
  message("The greeting is ", output$greeting) #You attempt to read from an output.

}

shinyApp(ui, server)
```



### Reacctive Programming

```{r}
ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)

server <- function(input, output, session) {
  output$greeting <- renderText({
    paste0("Hello ", input$name, "!")
  })
}
shinyApp(ui, server)
```

### Imperative Vs. Declarative Programming

**Imperative:** You issues a specific command ans iti's carried out immediately. (Like R script for data analysis)

**Declarative**: You express higher-level goals or describe important constraints, and rely on someone else to decide how and/or when to translate that into action.



### Laziness


```{r}
server <- function(input, output, session) {
  output$greting <- renderText({
    paste0("Hello ", input$name, "!")
  })
}
```

If you’re working on a Shiny app and you just can’t figure out why your code never gets run, double check that your UI and server functions are using the same identifiers.

### The reactive Graph

The reactive graph is a powerful tool for understanding how your app works. As your app gets more complicated, it’s often useful to make a quick high-level sketch of the reactive graph to remind you how all the pieces fit together. Throughout this book we’ll show you the reactive graph to help understand how the examples work, and later on, in Chapter 14, you’ll learn how to use reactlog which will draw the graph for you.

### Reactive expressions

Reactive expressions take inputs and produce outputs so they have a shape that combines features of both inputs and outputs. Hopefully, the shapes will help you remember how the components fit together.



```{r}

ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)
server <- function(input, output, session) {
  string <- reactive(paste0("Hello ", input$name, "!"))
  output$greeting <- renderText(string())
}
shinyApp(ui, server)
```


### Execution order

This concept is very important and different to most other R code, so I’ll say it again: the order in which reactive code is run is determined only by the reactive graph, not by its layout in the server function.

### Exercises

```{r}
ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)
server <- function(input, output, session) {
    greeting <- reactive(paste0("Hello ", input$name))
  output$greeting <- renderText(greeting())
}
shinyApp(ui, server)

```




## Reactive Expressions

This duality means we need some new vocab: We’ll use **producers** to refer to reactive inputs and expressions, and **consumers** to refer to reactive expressions and outputs. Figure 3.5 shows this relationship with a Venn diagram.


```{r}
library(ggplot2)

freqpoly <- function(x1, x2, binwidth = 0.1, xlim = c(-3, 3)) {
  df <- data.frame(
    x = c(x1, x2),
    g = c(rep("x1", length(x1)), rep("x2", length(x2)))
  )

  ggplot(df, aes(x, colour = g)) +
    geom_freqpoly(binwidth = binwidth, size = 1) +
    coord_cartesian(xlim = xlim)
}

t_test <- function(x1, x2) {
  test <- t.test(x1, x2)
  
  # use sprintf() to format t.test() results compactly
  sprintf(
    "p value: %0.3f\n[%0.2f, %0.2f]",
    test$p.value, test$conf.int[1], test$conf.int[2]
  )
}

# Simulated data
x1 <- rnorm(100, mean = 0, sd = 0.5)
x2 <- rnorm(200, mean = 0.15, sd = 0.9)

freqpoly(x1, x2)
cat(t_test(x1, x2))
#> p value: 0.009
#> [-0.42, -0.06]
```


The app (Inneficient version)

```{r}
ui <- fluidPage(
  fluidRow(
    column(4, 
      "Distribution 1",
      numericInput("n1", label = "n", value = 1000, min = 1),
      numericInput("mean1", label = "µ", value = 0, step = 0.1),
      numericInput("sd1", label = "σ", value = 0.5, min = 0.1, step = 0.1)
    ),
    column(4, 
      "Distribution 2",
      numericInput("n2", label = "n", value = 1000, min = 1),
      numericInput("mean2", label = "µ", value = 0, step = 0.1),
      numericInput("sd2", label = "σ", value = 0.5, min = 0.1, step = 0.1)
    ),
    column(4,
      "Frequency polygon",
      numericInput("binwidth", label = "Bin width", value = 0.1, step = 0.1),
      sliderInput("range", label = "range", value = c(-3, 3), min = -5, max = 5)
    )
  ),
  fluidRow(
    column(9, plotOutput("hist")),
    column(3, verbatimTextOutput("ttest"))
  )
)

server <- function(input, output, session) {
  output$hist <- renderPlot({
    x1 <- rnorm(input$n1, input$mean1, input$sd1)
    x2 <- rnorm(input$n2, input$mean2, input$sd2)
    
    freqpoly(x1, x2, binwidth = input$binwidth, xlim = input$range)
  }, res = 96)

  output$ttest <- renderText({
    x1 <- rnorm(input$n1, input$mean1, input$sd1)
    x2 <- rnorm(input$n2, input$mean2, input$sd2)
    
    t_test(x1, x2)
  })
}

shinyApp(ui, server)
```



The app (more efficient version)


```{r}
ui <- fluidPage(
  fluidRow(
    column(4, 
      "Distribution 1",
      numericInput("n1", label = "n", value = 1000, min = 1),
      numericInput("mean1", label = "µ", value = 0, step = 0.1),
      numericInput("sd1", label = "σ", value = 0.5, min = 0.1, step = 0.1)
    ),
    column(4, 
      "Distribution 2",
      numericInput("n2", label = "n", value = 1000, min = 1),
      numericInput("mean2", label = "µ", value = 0, step = 0.1),
      numericInput("sd2", label = "σ", value = 0.5, min = 0.1, step = 0.1)
    ),
    column(4,
      "Frequency polygon",
      numericInput("binwidth", label = "Bin width", value = 0.1, step = 0.1),
      sliderInput("range", label = "range", value = c(-3, 3), min = -5, max = 5)
    )
  ),
  fluidRow(
    column(9, plotOutput("hist")),
    column(3, verbatimTextOutput("ttest"))
  )
)


server <- function(input, output, session) {
  
  x1 <- reactive(rnorm(input$n1, input$mean1, input$sd1))
  x2 <- reactive(rnorm(input$n2, input$mean2, input$sd2))

  
  output$hist <- renderPlot({
    freqpoly(x1(), x2(), binwidth = input$binwidth, xlim = input$range)
  }, res = 96)

  output$ttest <- renderText({
    t_test(x1(), x2())
  })
}

shinyApp(ui, server)
```



You might be familiar with the “rule of three” of programming: whenever you copy and paste something three times, you should figure out how to reduce the duplication (typically by writing a function). This is important because it reduces the amount of duplication in your code, which makes it easier to understand, and easier to update as your requirements change.

In Shiny, however, I think you should consider the rule of one: whenever you copy and paste something once, you should consider extracting the repeated code out into a reactive expression. The rule is stricter for Shiny because reactive expressions don’t just make it easier for humans to understand the code, they also improve Shiny’s ability to efficiently rerun code.















